---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

##Things I learned on Jan. 27, 2022.

The command to set a working directory is setwd()

##Things I learned on Feb. 1, 2022

str_to_uppercase means the letters are uppercase

##Things I learned on Feb. 8, 2022

mutate(id = as.character(id))-- allows you to change the character type of a column. Key: change the "as.character" to "as.numeric" etc, etc. 

How to find columns where the value is NA-- filter(is.na(column_name)) %>%
How to find columns where the value is NOT NA-- filter(!is.na(column_name)) %>%

##Things I learned on Feb. 10, 2022
Standardize means to make a data set uniform e.g. make city names in "city" column all uppercase. 
str_detect(city,"^Anld") is a code that commands the program to detect situations where "Anld" exists in the data. 
Mutate is used to create a new variable. 
Order- ascending.... desc- descending

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```